---
title: "Economic Mobility and Bar Graphs"
author: "Laurie Heyer"
date: "1/26/2023"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```


## Make bar graph versions of a Du Bois data portrait

1. Explain how the following tribble represents occupation data in the [Du Bois data portrait](https://tile.loc.gov/storage-services/service/pnp/ppmsca/33800/33889v.jpg) that looks like two pie wedges.

2. From this data portrait, can you tell what percentage of domestic jobs are filled by White people and what percent are filled by Black people? Why or why not?

```{r}
# Make a data frame for the occupation visualization
occup <- tribble(
  ~sector,    ~race,    ~pct,
  "Agriculture", "black", 62, 
  "Agriculture", "white", 64,
  "Domestic", "black", 28,
  "Domestic", "white", 5.5,
  "Manufacturing", "black", 5,
  "Manufacturing", "white", 13.5,
  "Trade/Trans", "black", 4.5,
  "Trade/Trans", "white", 13,
  "Professions", "black", 0.5,
  "Professions", "white", 4
)


```

3. Use an appropriate geom to make at least two different bar graphs to visualize these data. Consider different choices for the x variable, and different choices of position. 

```{r}

```

## Read in Chetty data (with a twist)
```{r}
chetty <- read_excel("Lab1_Chetty_2014.xlsx", skip=1)
# The following code converts the variable top_100 from continuous to categorical
chetty <- chetty %>%
  mutate(top_100 = as.factor(top_100))
```

4. Which variables in the chetty data frame are appropriate x variables for a bar graph? 

5. Make two separate bar graphs for two different x variables. 

6. Make two more bar graphs that display proportions rather than counts of those same variables. 
```{r}

```

7. Make a bar graph that lets you compare the number of urban and rural CZ's in each of the four regions. 

8. Make a bar graph that lets you compare the total number of CZ's in each of the four regions, including the proporation that are urban and rural.

9. Make another bar graph that lets you easily compare the proportions of urban CZs across the four regions.
```{r}

```
10. Discuss the pros and cons of each visualization you created.

