---
title: "In-class activity, Thursday, 1-19"
author: "Laurie Heyer"
date: "1/19/2022"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

Read in the incarceration data

```{r}
prison <- read_excel("justice_NY_clean.xls",
                     skip=6)
```

1. Make a scatter plot of the % Under $25K vs rate. 
```{r}
ggplot(data = prison) + 
  geom_point(mapping = aes(x = pct_under_25K, y = rate))
```


2. Use aesthetics to indicate cost
```{r}
ggplot(data = prison) + 
  geom_point(mapping = aes(x = pct_under_25K, y = rate, size = cost))
```

3. Change the size, shape and/or color of all points in your scatterplot
```{r}
ggplot(data = prison) + 
  geom_point(mapping = aes(x = pct_under_25K, y = rate), 
             shape = 24, size = 5, color = "black", fill = "purple")
```

4. What problem(s) do you encounter if you try to use the variable $cost original instead of cost?

* One problem is that $cost original is not a standard variable name, so you have to enclose it in backticks (not apostrophes!)
* Another problem is $cost original is not a numeric variable, so the scale makes no sense. Each individual value in the data frame gets its own dot size, and they are not in numberical order. 
```{r}
ggplot(data = prison) + 
  geom_point(mapping = aes(x = pct_under_25K, y = rate, size = `$cost original`))
```

# Try knitting this to Word, html and pdf

## Do they all work?

### Compare the formatting of this line and the two above

* a bullet point -- often need a blank line above a list to get it to format correctly when you knit
- same level, with minus sign
  + one tab with plus
    * two tabs, back to star
      + three tabs, a plus

  
1. numbered list
    a. two tabs
        i. four tabs, roman numeral
    a. back to two with another a
    a. yet another a
1. second item
